Faithful Variable Screening for High-Dimensional Convex Regression

### 2014-07-01 tex/scam/introduction.tex
restructured related work

Variable selection for nonparametric
regression under smoothness constraints is difficult without
making additional strong assumptions. 
\cite{lafferty2008rodeo} show that computationally
efficient, near minimax-optimal estimation is possible, but in ambient
dimensions that scale only as $p = O(\log n)$; see also \cite{BertinLecue}.  This is in contrast
to the $p=O\bigl(e^{n^c}\bigr)$ scaling enjoyed by sparse linear
models \cite{Wain:09a}. \citet{dalalyan:12} achieve exponential scaling $p=O(e^n)$
under certain Fourier smoothness conditions, but require
that the number of relevant variables $s$ must be less than $\log n$.

Approximating the regression function by a sum of one-dimensional
functions, known as sparse additive models \citep{Ravikumar:09}, is a
practical alternative to fully nonparametric function estimation.  
Sparse additive models have been further studied in by \cite{Meier09},
and by \cite{Raskutti:12}.  But 
the additive assumption is limited.  In particular, the natural idea
of first selecting the single variable effects, then the pairwise
effects, and so on, does not in general lead to consistent variable
selection.  In other words, the general nonparametric model is not
additively faithful.  Remarkably, the additional assumption of
convexity does lead to consistent variable selection, as we show
here. In addition, we show that the high dimensional scaling $n =
O\big(\textrm{poly}(s) \log p\big)$ is achievable for sparse convex
additive models. Thus, with respect to variable selection, the
geometric convexity constraint is quite different from the smoothness
constraints imposed in traditional nonparametric regression.

A key to our approach is the observation that least squares
nonparametric estimation under convexity constraints is equivalent to
a finite dimensional quadratic program.  Specifically, the infinite
dimensional optimization 
\begin{align}
\begin{split}
\text{minimize} & \quad \sum_{i=1}^n (Y_i - m(x_i))^2 \\
\text{subject to} &  \quad m:\reals^p\rightarrow\reals\ \text{is
  convex}
\end{split}
\end{align}
is precisely equivalent to the finite dimensional quadratic
program 
\begin{align}
\begin{split}
\label{eq:outer}
\text{minimize}_{h, \beta} & \;\; \sum_{i=1}^n (Y_i - h_i)^2 \\
\text{subject to} & \;\; h_j \geq h_i + \beta_i^T (x_j-x_i),\; \text{for
    all $i,j$}.
\end{split}
\end{align}
%\end{equation}
%See \cite{Boyd04}, Section 6.5.5.
Here $h_i$ is the estimated function value $m(x_i)$, and the vectors
$\beta_i \in \reals^d$ represent supporting hyperplanes to the
epigraph of $m$.  Importantly, this finite dimensional quadratic program does
not have tuning parameters for smoothing the function. Such parameters are the bane
of nonparametric estimation.

Estimation of convex functions arises naturally in several
applications.  Examples include geometric programming \cite{Boyd04},
computed tomography \cite{Prince:90}, target reconstruction
\cite{Lele:92}, image analysis \cite{Golden:06} and circuit design
\cite{Hannah:12}.  Other applications include queuing theory
\cite{Chen:01} and economics, where it is of interest to estimate
concave utility functions \cite{Pratt:68}.  See \cite{Lim:12} for
other applications.

Beyond cases where the assumption of convexity is
natural, we offer that the convexity assumption is attractive as a
tractable, nonparamametric relaxation of the linear model.  In
addition to the lack of tuning parameters, other than the
regularization parameter $\lambda$ to control the level of sparsity,
the global convexity assumption leads to effective, scalable algorithms.  We
demonstrate use of our approach on experiments with standard
regression data sets, in a comparison with sparse linear models
(lasso).

%In the following section we give a high-level summary of our technical
%results, including additive faithfulness, variable selection 
%consistency, and high dimensional scaling.  In Section~X...

\def\mathbf#1{\mbox{\boldmath $#1$}}

>>>>

Estimation of convex functions arises naturally in several
applications.  Examples include geometric programming \citep{Boyd04},
computed tomography \citep{Prince:90}, target reconstruction
\citep{Lele:92}, image analysis \citep{Golden:06} and circuit design
\citep{Hannah:12}.  Other applications include queuing theory
\citep{Chen:01} and economics, where it is of interest to estimate
concave utility functions \citep{Pratt:68}.  See \cite{Lim:12} for
other applications.  
Beyond cases where the assumption of convexity is
natural, the convexity assumption can be attractive as a
tractable, nonparamametric relaxation of the linear model.

Variable selection in general nonparametric regression or function
estimation is a notoriously difficult problem. \citet{lafferty2008rodeo} develop a greedy procedure for
adjusting bandwidths in a local linear regression estimator,
and show that the procedure achieves the minimax rate
as if the relevant variables were isolated in advance.
But the method only provably scales to dimensions $p$ that 
grow logarithmically in the sample size $n$, so that $p = O(\log n)$.  Note that this
is the opposite of the high dimensional scaling behavior
known to hold for sparsity selection in linear models
using $\ell_1$ penalization, where the sample size $n$
is logarithmic in the dimension $p$. \citet{bertin:08}
develop an optimization-based approach in
the nonparametric setting, applying the lasso
in a local linear model at each test point.  Here again,
however, the method only scales as $p = O(\log n)$,
the low-dimensional regime.
An approximation theory approach to the same
problem is presented in \cite{devore:11}, 
using techniques based on hierarchical hashing schemes,
similar to those used for ``junta'' problems \cite{mossel:04}.
Here it is shown that the sample complexity scales as $n > \log p$ 
if one adaptively selects the points on
which the high-dimensional function is evaluated.

\citet{dalalyan:12} show that the exponential scaling $p=O(\log n)$ is achievable if the underlying function is assumed to be smooth with respect to a Fourier basis. They also give
support for the intrinsic difficulty of variable
selection in nonparametric regression, giving lower bounds 
showing that sparsistency is not possible if $n < \log p$ or if $n <
\exp s$, where $s$ is the number of relevant variables.
Variable selection over kernel classes is studied
by \citet{Kolch:10}.

Perhaps most closely related to the present work, is the framework
studied by \cite{Raskutti:12} for sparse additive models, where sparse
regression is considered under an additive assumption, with each
component function belonging to an RKHS.  An advantage of working over
an RKHS, in contrast to the other papers mentioned above, is that
nonparametric regression with a sparsity-inducing regularization penalty can be
formulated as a finite dimensional convex cone optimization.
On the other hand, 
smoothing parameters for the component Hilbert spaces
must be chosen, leading to extra tuning parameters
that are difficult to select in practice.  In addition, 
the additive model must be assumed to
be correct for sparsistent variable selection.

An attraction of the convex function estimation framework we consider
in this paper is that the additive model can be used for convenience,
without assuming it to actually hold.  While nonparametric, the
problem is naturally formulated using finite dimensional convex
optimization, but with no additional tuning parameters.  As we show
below, our method scales to high dimensions, with a dependence on the
intrinsic dimension $s$ that scales polynomially, rather than
exponentially as in the general case analyzed in \cite{dalalyan:12}.

%* Laetitia Comminges, Arnak S. Dalalyan 
%Tight conditions for consistency of variable selection 
%In the context of high dimensionality. 
%Ann. Statist, 40(5), 2667-2696, 2012 
%
%Comment: They estimate the coefficient of the regression function with respect to a Fourier basis and then threshold the estimated coefficients. They prove that sparsistency is achievable so long as $n > log p$ and $n > exp(s)$. They assume that the functions are smooth with respect to the Fourier basis. The fourier coefficients however seem difficult to estimate; it appears that the knowledge of the true density is required. They also provide 
%
%
%* Bertin, K. and Lecu´e, G. (2008). Selection of variables 
%and dimension reduction in high-dimensional non-parametric 
%regression. Electron. J. Stat. 2 1224–1241 
%
%Comment: They assume Holder smoothness and use multiple L1 regularized local polynomial smoothing to detect the sparsity pattern. They require that $p < log n$.
%
%
%* "Approximation of Functions of Few Variables in High Dimensions" 
%by DeVore, Petrova and Wojtaszczyk. 
%
%Comment: They show that variable selection under a scaling of
%

%Variable selection for nonparametric
%regression under smoothness constraints is difficult without
%making additional strong assumptions. 
%\cite{lafferty2008rodeo} show that computationally
%efficient, near minimax-optimal estimation is possible, but in ambient
%dimensions that scale only as $p = O(\log n)$; see also \cite{BertinLecue}.  This is in contrast
%to the $p=O\bigl(e^{n^c}\bigr)$ scaling enjoyed by sparse linear
%models \cite{Wain:09a}. \citet{dalalyan:12} achieve exponential scaling $p=O(e^n)$
%under certain Fourier smoothness conditions, but require
%that the number of relevant variables $s$ must be less than $\log n$.
%
%Approximating the regression function by a sum of one-dimensional
%functions, known as sparse additive models \citep{Ravikumar:09}, is a
%practical alternative to fully nonparametric function estimation.  
%Sparse additive models have been further studied in by \cite{Meier09},
%and by \cite{Raskutti:12}.  But 
%the additive assumption is limited.  In particular, the natural idea
%of first selecting the single variable effects, then the pairwise
%effects, and so on, does not in general lead to consistent variable
%selection.  In other words, the general nonparametric model is not
%additively faithful.  Remarkably, the additional assumption of
%convexity does lead to consistent variable selection, as we show
%here. In addition, we show that the high dimensional scaling $n =
%O\big(\textrm{poly}(s) \log p\big)$ is achievable for sparse convex
%additive models. Thus, with respect to variable selection, the
%geometric convexity constraint is quite different from the smoothness
%constraints imposed in traditional nonparametric regression.
%

In the following section we give a high-level summary of our technical
results, including additive faithfulness, variable selection 
consistency, and high dimensional scaling.  In Section~X...
